Object Capability Patterns:\\Policies and Specifications

### 2017-08-17 article.tex
cooperation

\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Path Condition 2 (NPC2)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$]$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
\ablock \loimplies \linebreak
	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
\hrule
\end{logic}
, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use the condition to expand the condition till we reach the terminating o'. \textsc{npc2} gives us the present path configurations that will allow an eventual path from o to o' as a necessary condition for o to modify o'.But if objects are path-isolated, how do we allow them to cooperate?

@@

The logics we have developed show that security of an object can be enforced in an OCap model by:
\begin{itemize}
\item being careful about our present path configurations so that we can ensure eventual path-isolation between the object and untrusted object
\item specifying security policies on how \textit{all} objects with the direct capability of our trusted object may modify the object
\end{itemize}
Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authoriy \textit{cannot} enforce security through path isolation, which is a distinctive feature of OCap models. How do we then expect objects with the direct capability of o' may modify o' through some security policy? The nice thing is we already have a predicate MayAffect$^{Eve}$(*,*) that can accomodate these security policies. Notice how in \textsc{NEPC*}, object Y* sits as the last object in the path from o to o' and has the capability of o'. This gives us a clue as to how we can enable protection of o' without denying paths and writing security policies that allow

\subsection{What about Cooperation?}
At this juncture, it is good to know we can enforce protection through denying possible path configurations, but in practice this is still not useful. Cooperation between objects require paths to exist between objects so it is rare or almost impossible to have two objects o and o' that cooperate while \loneg MayAccess$^{Ind,Eve}$(o,o') holds. The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o' while \textit{still} allowing paths to exist between them. This is possible, because of Rule 3 that says objects may have private encapsulation of state and behaviour. To begin, let us assume a simple object o' that has \textit{only} private behaviours. In such a scenario, since all fields in our system are also assumed to be private, then we know:
\begin{logic} \loturns\ \loforall X*\loin Obj, X*\loneq o'. \loneg MayAffect(X*,o') \end{logic}
Notice, how protection of o' can occur independent of paths and capability, due to object encapsulation. This means that objects can specify security policies that it chooses to enforce.

@@

If we have some starting configuration in a system where there is only one object o that has the capability of o', and no other objects have the capability of o', then we know any eventual path to o' has to include and end with the sub-path from o to o', \textit{unless} o leaks the capability of o' to some other object X*. Notice how the sub-path o to o' is also a direct capability from o to o'. Object o, therefore, has the power to dictate the policies that can affect o' through itself, and state more precise stronger necessary conditions for the MayAffect$^{Eve}${(o,o') predicate, conditions other than path existence which we have argued. To put it in another way, we can now allow a path to exist from o to o', \textit{and} still enforce \loneg MayAffect$^{Eve}$(o,o'), because the negation of these stronger security policy conditions results in weaker antecedents for protection consequence.\\

We now see the power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another. The logics we have developed so far illuminate the power of attenuating objects X* that can enable protection. Indeed attenuating objects feature prominently in the literature of OCap patterns which we shall see in the next section.

@@

\begin{logic}[NodeVul]
\hrule\null
\textsc{\normalsize *Node Vulnerability}\\
\loforall o, n. MayAccess$^{Dir,Now}$(o,n)\linebreak \ablock\ablock \loimplies \linebreak
\ablock \ablock MayAffect$^{Now}$(o,n) \loand \linebreak
\ablock \ablock \loforall n'. MayAccess$^{Dir,Now}$(o,n')
\end{logic}
The vulnerability of a node lies in the fact that it contains a public method \texttt{setProperty(key,value)} that will modify an internal mapping data structure. A node also has a public field \texttt{parent} that will return the capability of its parent node. Consequently, this allows an object which has the capability of any one node in the tree to navigate up to the root node (Document), and consequently navigate to all other nodes in the tree.
%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\small
\bibliographystyle{plain}
\bibliography{ref}

>>>>

\begin{logic}
\hrule\null
\textsc{\normalsize *Necessary Path Condition 2 (NPC2)}\\
M,\losigma\ \loturns\ MayAffect$^{Eve}$(o,o')\linebreak
\ablock \loimplies\linebreak
\ablock \loexists X*\loin Obj.$[($MayAccess$^{Ind,Now}$(o,X*) \loor \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Dir,Now}$(X*,o)$)$\linebreak
\ablock \ablock \ablock \ablock \loand \linebreak
\ablock \ablock \ablock \quad MayAccess$^{Ind,Eve}$(X*,o')$]$
\linebreak \\
\textit{By contraposition,}\linebreak
M,\losigma\ \loturns\ 
\loforall X*\loin Obj. $[$\loneg MayAccess$^{Ind,Now}$(o,X*) \loand \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Dir,Now}$(X*,o)$]$\linebreak
\ablock \ablock \ablock \ablock \loor \linebreak
\ablock \ablock \ablock \ablock \loneg MayAccess$^{Ind,Eve}$(X*,o')$]$\linebreak
\ablock \loimplies \linebreak
	\ablock\loneg MayAffect$^{Eve}(o,o')$\linebreak
\hrule
\end{logic}
, where we can always use \textsc{gpec} to expand indirect paths into direct capabilities till we reach the terminating origin o and recursively use the condition to expand the condition till we reach the terminating o'. \textsc{npc2} gives us the present path configurations that will allow an eventual path from o to o' as a necessary condition for o to modify o'. But if objects are path-isolated, how do we allow them to cooperate?

@@

The logics we have developed show that security of an object can be enforced in an OCap model by:
\begin{itemize}
\item being careful about our present path configurations so that we can ensure eventual path-isolation between the object and untrusted object
\item specifying security policies on how \textit{all} objects with the direct capability of our trusted object may modify the object
\end{itemize}
Therein lies also the difference between OCap and non-OCap models, because non-OCap models that allow forging of capabilities or a global ambient authoriy \textit{cannot} enforce security through path isolation, which is a distinctive feature of OCap models.\\

The similarity is that both models allow us to specify policies \textit{within} objects as to whether an object's behaviour can be called (although non OCap systems can do so 'externally' through a global ambient authority). 'Internal' object protection can be done through using defensive programming techniques like encapsulation and checking of conditions for behaviours to succeed. In a way these forms of protection can be a form of 'stack-based' access control from the called object's perspective, because they only happen \textit{after} the calling object already has already called the object.\\

How do we then these internal object security policies in our logical framework? The nice thing is we already have a predicate MayAffect$^{Eve}$(*,*) that can accomodate this. Notice how in \textsc{NEPC2*}, object Y* sits as the last object in the path from o to o' \textit{and} has the capability of o'. This gives us a clue as to how we can enable protection of o' without denying paths, by \textit{specifying whether Y* can affect o' explicitly in security policies} that are independent of paths.

\subsection{What about Cooperation?}
It is good to know we can enforce protection through denying possible path configurations, but in practice this is still not useful. Cooperation between objects require paths to exist between objects so it is rare or almost impossible to have two objects o and o' that cooperate while \loneg MayAccess$^{Ind,Eve}$(o,o') holds. The quintessential question in OCap systems is whether we can enforce protection such as controlling whether o may affect o' while \textit{still} allowing paths to exist between them. This is possible, because of Rule 3 that says objects may have private encapsulation of state and behaviour. To begin, let us assume a simple object o' that has \textit{only} private behaviours. In such a scenario, since all fields in our system are also assumed to be private, then we know:
\begin{logic} \loturns\ \loforall X*\loin Obj, X*\loneq o'. \loneg MayAffect(X*,o') \end{logic}
Notice, how protection of o' can occur independent of paths and capability, due to object encapsulation. This means that objects can specify security policies that it chooses to enforce.

@@

If we have some starting configuration in a system where there is only one object o that has the capability of o', and no other objects have the capability of o', then we know any eventual path to o' has to include and end with the sub-path from o to o', \textit{unless} o leaks the capability of o' to some other object. Notice how the sub-path o to o' is also a direct capability from o to o'. Object o, therefore, has the power to dictate the policies that can affect o' through itself (since it holds the direct capability of o'), and state more precise stronger necessary conditions for the MayAffect$^{Eve}${(o,o') predicate. To put it in another way, we can now allow a path to exist from o to o', \textit{and} still enforce \loneg MayAffect$^{Eve}$(o,o'), because the negation of these stronger security policy conditions results in weaker antecedents for \loneg MayAffect$^{Eve}$(o,o'). This can be done simply, be ensuring that object o has the capability of o' stored in some private field but has no behaviours that will make use and call this capability.\\

A key result from this analysis shows the importance of attenuating objects, objects that sit as the last line of defence in the eventual paths to our protected objects. In the OCap pattern examples we will go through in the next section, protection of an object is done through some attenuating object that the trusted object knows will never leak its capability directly (the trusted object can ensure this if it creates those attenuating objects), and that protection is enforced through some security policies within those attenuating objects.\\\

If the protected object can guarantee through the system of configuration that those attenuating objects are the \textit{only} objects in the entire system that holds its capability in \textit{all} eventual states, then the protected object can dictate the conditions and degree of control on how it can be used by untrusted objects in all eventual states through those attenuating objects. This ability of the protected object to dictate control across time and conditions enables cooperation, where conditions on how the protected object will cooperate is written in those attenuating objects.
\\

We now see the power of OCap patterns hence lies in providing concrete examples of a system of cooperation that allows the existence of paths between objects for cooperation while still dictating the degree of control of one object can have over another.

@@

\begin{logic}[NodeVul]
\hrule\null
\textsc{\normalsize *Node Vulnerability}\\
\loforall o, n. MayAccess$^{Dir,Now}$(o,n)\linebreak \ablock\ablock \loimplies \linebreak
\ablock \ablock MayAffect$^{Now}$(o,n) \loand \linebreak
\ablock \ablock \loforall n'. MayAccess$^{Dir,Now}$(o,n')
\end{logic}
The vulnerability of a node lies in the fact that it contains a public method \texttt{setProperty(key,value)} that will modify an internal mapping data structure. A node also has a public field \texttt{parent} that will return the capability of its parent node. Consequently, this allows an object which has the capability of any one node in the tree to navigate up to the root node (Document), and consequently navigate to all other nodes in the tree.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\small
\bibliographystyle{plain}
\bibliography{ref}
